---
title: "Midterm 1 W24"
author: "Hugo Mahatdejkul"
date: "`r Sys.Date()`"
output:
  html_document: 
    keep_md: yes
  pdf_document: default
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your code must be organized, clean, and run free from errors. Remember, you must remove the `#` for any included code chunks to run. Be sure to add your name to the author header above. 

Your code must knit in order to be considered. If you are stuck and cannot answer a question, then comment out your code and knit the document. You may use your notes, labs, and homework to help you complete this exam. Do not use any other resources- including AI assistance.  

Don't forget to answer any questions that are asked in the prompt!  

Be sure to push your completed midterm to your repository. This exam is worth 30 points.  

## Background
In the data folder, you will find data related to a study on wolf mortality collected by the National Park Service. You should start by reading the `README_NPSwolfdata.pdf` file. This will provide an abstract of the study and an explanation of variables.  

The data are from: Cassidy, Kira et al. (2022). Gray wolf packs and human-caused wolf mortality. [Dryad](https://doi.org/10.5061/dryad.mkkwh713f). 

## Load the libraries
```{r message=FALSE, warning=FALSE}
library("tidyverse")
library("janitor")
```

## Load the wolves data
In these data, the authors used `NULL` to represent missing values. I am correcting this for you below and using `janitor` to clean the column names.
```{r message=FALSE, warning=FALSE}
wolves <- read.csv("data/NPS_wolfmortalitydata.csv", na = c("NULL")) %>% clean_names()
```

## Questions
Problem 1. (1 point) Let's start with some data exploration. What are the variable (column) names?  
```{r}
names(wolves)
#"park"         "biolyr"       "pack"         "packcode"     "packsize_aug" "mort_yn"      "mort_all"     "mort_lead"    "mort_nonlead"
# "reprody1"     "persisty1" are the names.
```

Problem 2. (1 point) Use the function of your choice to summarize the data and get an idea of its structure.  
```{r}
glimpse(wolves)
#summary(wolves)
#dim(wolves)
```

Problem 3. (3 points) Which parks/ reserves are represented in the data? Don't just use the abstract, pull this information from the data.  
```{r}
wolves %>%
  group_by(park) %>%
  count(park)

#DENA, GNTP, VNP,YNP, YUCH
```


Problem 4. (4 points) Which park has the largest number of wolf packs?
```{r}
wolves %>%
  group_by(park) %>%
  tabyl(park)

#DENA has the largest number of wolf packs, with 340 total wolf packs!
```
```{r}
wolves %>%
  group_by(park) %>%
  summarize(n=n_distinct(pack))
```



Problem 5. (4 points) Which park has the highest total number of human-caused mortalities `mort_all`?
```{r}
wolves %>%
  group_by(park) %>%
  summarize(mortalities=sum(mort_all)) %>%
  arrange(desc(mortalities))
#YUCH park has the highest total number of human caused moralities
```

The wolves in [Yellowstone National Park](https://www.nps.gov/yell/learn/nature/wolf-restoration.htm) are an incredible conservation success story. Let's focus our attention on this park.  

Problem 6. (2 points) Create a new object "ynp" that only includes the data from Yellowstone National Park.  
```{r}
ynp<-
  wolves %>%
  filter(park=="YNP")
ynp
```

Problem 7. (3 points) Among the Yellowstone wolf packs, the [Druid Peak Pack](https://www.pbs.org/wnet/nature/in-the-valley-of-the-wolves-the-druid-wolf-pack-story/209/) is one of most famous. What was the average pack size of this pack for the years represented in the data?
```{r}
ynp%>%
  filter(pack=="druid") %>%
  summarize(mean_pack_size_druid=mean(packsize_aug))
#Avg packsize for druids were 13.933 over the years in the data

```

Problem 8. (4 points) Pack dynamics can be hard to predict- even for strong packs like the Druid Peak pack. At which year did the Druid Peak pack have the largest pack size? What do you think happened in 2010?
```{r}
ynp %>%
  filter(pack=="druid") %>%
  group_by(biolyr) %>%
  count(packsize_aug) %>%
  arrange(desc(packsize_aug))
#2001 had the largest pack size. In 2010, the packsize was at a big fat 0. I hear Yosemete has a lot of troubles with fires during drought seasons, so it is possible that there was a drought in California and then fires led to the decrease in druid wolf packs.

```

Problem 9. (5 points) Among the YNP wolf packs, which one has had the highest overall persistence `persisty1` for the years represented in the data? Look this pack up online and tell me what is unique about its behavior- specifically, what prey animals does this pack specialize on?  
```{r}
ynp %>%
  select(biolyr,pack,persisty1) %>%
  group_by(pack) %>%
  summarize(overallp_persistence_by_pack=sum(persisty1)) %>%
  arrange(desc(overallp_persistence_by_pack))
#mollies had the overall highest persistence. This animal specializes in hunting bison. Unlike some other wolf packs that primarily feed on elk, the Mollie pack has become skilled in preying on Bison. They use the snow to their advantage to make what would be an extremely difficult hunt, to a manageable one. Snow tend to weaken the bison and provide Mollie's pack with the perfect opportunity to attack.
```

Problem 10. (3 points) Perform one analysis or exploration of your choice on the `wolves` data. Your answer needs to include at least two lines of code and not be a summary function.  
```{r}
wolves %>%
  select(biolyr,packsize_aug) %>%
  filter(packsize_aug>=10) %>%
  group_by(biolyr) %>%
  summarize(mean_packsize_by_year=mean(packsize_aug)) %>%
  arrange(desc(mean_packsize_by_year))
#filtering out all the smaller packs, I wanted to see which year mounted the largest pack sizes on average. I expected the year covid hit to have the largest pack size since there would be less human interaction with the animals, and it seems that was the case.
```


#save knitted file as pdf and then upload it into gradescope thru canvas