---
title: "BIS15L_AI_hmwk"
author: "Hugo Mahatdejkul"
date: "`r Sys.Date()`"
output: 
  html_document: 
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
This homework is a bit less traditional this week. There are some errors in the code below. In addition add notes and titles to the code chunks that explains what the code chunck does. Use AI tools, notes from class, or any other resources to help fix the errors. Feel free to edit any plots to your preference. Don't forget to answer the two questions at the bottom. Upon completion, you should not have to '#' out any of the code chunks.

Push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.   

Again make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  


```{r}
library(tidyverse)
library(janitor)
library(dplyr)
```


1.
```{r}
blood_count <- read_csv("data/blood_count_dataset.csv")
```

2. 
```{r}
glimpse(blood_count)

```
#fix this code
3. 
```{r}
blood_count %>%
  group_by(Gender) %>%
  summarize(min = min(Red_Blood_Cells),
            mean = mean(Red_Blood_Cells), 
            max = max(Red_Blood_Cells))
```


4. 
```{r}
options(scipen = 999, digits = 2)

blood_count %>% 
  filter(Gender == "Male") %>%
  select(3:9) %>%
  colMeans(na.rm = TRUE)

```

5. 
```{r}
blood_count %>% 
  filter(Gender == "Female") %>%
  select(3:9) %>%
  colMeans(na.rm = T)
```


6. 
```{r}
blood_count %>%
  pivot_longer(cols = c(3:9), names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Gender, y = Values, fill = Gender)) +
  geom_boxplot() +
  facet_wrap(~ Test, scales = "free") + 
  labs(title = "Boxplot of Blood Count Tests",
       x = "Test",
       y = "Values")
```

7. 
```{r}
grouped_bc <- blood_count %>%
  mutate(Age_Group = case_when(
    Age >= 20 & Age <= 29 ~ "20s",
    Age >= 30 & Age <= 39 ~ "30s",
    Age >= 40 & Age <= 49 ~ "40s",
    Age >= 50 & Age <= 59 ~ "50s",
    Age >= 60 & Age <= 70 ~ "60s")) # there were only 4 people in 70s, so I combined it with 60s

head(grouped_bc)

```

8. 
```{r}
grouped_bc %>%
  pivot_longer(cols = c(3:9), names_to = "Test", values_to = "Values") %>%
  ggplot(aes(x = Age_Group, y = Values, fill = Age_Group)) +  
  geom_boxplot() +
  facet_wrap(~ Test, scales = "free") +  
  labs(title = "Boxplot of Blood Count Tests by Age Group",
       x = "Age Group",
       y = "Values")

```
9. What are two drawbacks and two benefits of AI?
I think the biggest drawbacks of AI is that it doesn't always know what I am asking and thus sometimes the solution it offers makes absolutely no sense what soever. Another drawback is that it can be very time consuming to train the AI to do what you want it to do. The benefits of AI is that it can be very helpful in automating tasks that would take a long time to do manually. Another benefit is that it can be very helpful in finding patterns in data that would be very difficult to find manually.
10. Do you think you will use AI for coding in the future? Why or why not?
Defintely! I think that AI can be very helpful in doing simple tasks like creating apps where there is a base template we need to be filled out. The Shinyapp has the 3 main chunks in the code and I think AI can be helpful in constructing the UI chunk. I have been tinkering with it for the final project and it has helped out a lot

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences. 