---
title: "BIS15L_AI_class"
author: "Your Name"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Overview
Today, we're exploring how AI can aid in writing, comprehending and troubleshooting our R code. AI is a very valuable tool, but it's important not to rely on it entirely. Building a strong R foundation is still important. These AI tools are not perfect and can be inaccurate solutions. In addition these tools can't read our minds, so understanding R helps us generate better prompts to get better responses. In short, AI should be used as a tool, not a crutch. Additionally, do not forget about other useful resources like Stack Overflow/Exchange, online tutorials, and GitHub issues pages for learning and troubleshooting. Let's combine AI with traditional methods to master R effectively.

This overview isn't exhaustive of all possible AI tools for R coding, but I believe it is a good start and covers everything you need.


**Make sure to enable Github Copilot in RStudio.**
Go to Tools -> Global Options -> Copilot -> Enable Git

## Learning Goals: 
*At the end of this exercise, you will be able to:*  
1. Explain what AI is and what AI tools can be used for R.
2. Enable Github Copilot in RStudio.
3. Use Copilot/ChatGPT to write code in RStudio.
4. Use Copilot/ChatGPT for troubleshooting code in RStudio.


## Exploration Questions:
Lets get some get a feel for what copilot (or ChatGPT) does, use Github Copilot to autocomplete the questions below. You can alternatively paste the questions in ChatGPT or Microsoft Copilot.
1. What is AI?
AI, or Artificial Intelligence, refers to the simulation of human intelligence processes by machines, especially computer systems. These processes include learning (the acquisition of information and rules for using the information), reasoning (using rules to reach approximate or definite conclusions), and self-correction. AI can be categorized into narrow AI, which is designed for a specific task, and general AI, which aims to mimic human intelligence across a wide range of tasks.
2. What is ChatGPT?
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It assists developers by providing context-aware code suggestions and autocompletions as they write code within their Integrated Development Environment (IDE). GitHub Copilot is trained on a vast amount of publicly available code and programming documentation, allowing it to generate code snippets, function calls, and even entire functions based on the context of the code being written. 
3. What is Github Copilot?
GitHub Copilot is an AI-powered code completion tool developed by GitHub in collaboration with OpenAI. It assists developers by providing context-aware code suggestions and autocompletions as they write code within their Integrated Development Environment (IDE). Copilot is trained on a vast dataset of code from public repositories and programming d
ocumentation, allowing it to generate code snippets, function calls, and even entire functions based on the context of the code being written. It supports multiple programming languages and aims to enhance productivity by offering intelligent suggestions and reducing the need for manual coding.
4. [Try it yourself: Insert your own question here]

## Load the libraries  
```{r message=FALSE, warning=FALSE}
library(tidyverse)
library(janitor)
library(dplyr)
library(shiny)
library(shinydashboard)
```

## Load the Data: 
```{r}
#load the Netflix Data in your Data folder, save it under the object name netflix
netflix<-read_csv("data/netflix_titles.csv")

```

*Did copilot help you write the code above? Could anything about the prompt be improved to help Copilot understand what you want to do more accurately?*


## Lets Practice:
Explore the elements in your data: include a comment within your code chunk to prompt Copilot to use the function of your choice
```{r}
head(netflix)

```
*Did copilot use your preferred method to get the summary of the data? If not, how could you improve the prompt to get the desired output?*
```{r}
names(netflix)
```


Sort out the movies from the netflix into an new object called netflix_movies
```{r} 
netflix_movies <-
  netflix %>%
  filter(type=="Movie")
```


Which director has the most movies on netflix?
```{r}
netflix_movies %>%
#  filter(director!=("NA")) %%
  group_by(director) %>%
  summarize(director_count=n()) %>%
  arrange(desc(director_count))
```
  ```
*Do you see anything wrong with the table above, look at row 1. Write a prompt below to fix this.*

[insert your prompt here]
```{r}

```



Which director had the most movies by release year and remove the NA's?
```{r}


```
*See anything else interesting about this table, look at the first director on the table. Could this be a problem and how can we edit the dataset to fix this? Try writing a prompt.*


[Insert your prompt here]
```{r}

```



Make a bar plot of the top directors with the most movies on netflix
```{r}


```
*Do you see any issues with this plot? How could we improve the prompt to get a better boxplot? Try writing a prompt.*

[Insert your prompt here]
```{r}

```

*How about for just 2020*



[Insert your prompt here]
```{r}

```


*Take a few minutes to make your own plot(s) using prompts, edit prompts until you get a plot that you like.*






## Lets make the top 10 directors barplot into a Shiny App: use copilot or Chatgpt


*Now use ChatGPT, Copilot or previous notes to fix the code*


```{r}

# Sample data
data <- data.frame(
  plant_type = c("tree", "fruit", "vegetable", "flower"),
  height = c(10, 5, 7, 8),
  weight = c(20, 15, 18, 12)
)

# Define UI
ui <- fluidPage(
  titlePanel("Plant Data Visualization"),
  
  sidebarLayout(
    sidebarPanel(
      checkboxGroupInput("variables",
                         label = h3("Variables to Display"),
                         choices = list("Height" = "height",
                                        "Weight" = "weight"),
                         selected = c("height", "weight")
      )
    ),
    
    mainPanel(
      plotOutput("plot")
    )
  )
)

# Define server logic
server <- function(input, output) {
  
  output$plot <- renderPlot({
    # Filter data based on selected variables
    plot_data <- data[, c("plant_type", input$variables)]
    
    # Melt data for plotting
    plot_data <- melt(plot_data, id.vars = "plant_type")
    
    # Plot
    ggplot(plot_data, aes(x = plant_type, y = value, color = variable)) +
      geom_point() +
      labs(x = "Plant Type", y = "Value", color = "Variable") +
      theme_minimal()
  })
}

# Run the application
shinyApp(ui = ui, server = server)

```





## That's it! Let's take a break and then move on to working on group projects.
