---
title: "Lab 4 Homework"
author: "Hugo Mahatdejkul"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

## Data
For the homework, we will use data about vertebrate home range sizes. The data are in the class folder, but the reference is below.  

**Database of vertebrate home range sizes.**  
Reference: Tamburello N, Cote IM, Dulvy NK (2015) Energy and the scaling of animal space use. The American Naturalist 186(2):196-211. http://dx.doi.org/10.1086/682070.  
Data: http://datadryad.org/resource/doi:10.5061/dryad.q5j65/1  

**1. Load the data into a new object called `homerange`.**
```{r}
getwd()

homerange<-read.csv("Tamburelloetal_HomeRangeDatabase.csv")

```

**2. Explore the data. Show the dimensions, column names, classes for each variable, and a statistical summary. Keep these as separate code chunks.**  
```{r}
dim(homerange)
```

```{r}
names(homerange)
```

```{r}
glimpse(homerange)
```

```{r}
summary(homerange)
```

**3. Change the class of the variables `taxon` and `order` to factors and display their levels.**  
```{r}
homerange$taxon<-as.factor(homerange$taxon)
class(homerange$taxon)
levels(homerange$taxon)
```

```{r}
homerange$order<-as.factor(homerange$order)
class(homerange$order)
levels(homerange$order)
```

**4. What taxa are represented in the `homerange` data frame? Make a new data frame `taxa` that is restricted to taxon, common name, class, order, family, genus, species.**  
```{r}
select(homerange,taxon)
#Mammals, turtles,tortoises,lizards,snakes, birds,river fish,lake fish, marine fish
```
Mammals, turtles,tortoises,lizards,snakes, birds,river fish,lake fish, marine fish

```{r}
names(homerange) #taxon, common name, class, order, family, genus, species
taxa<-select(homerange,taxon:species)
taxa
```

**5. The variable `taxon` identifies the common name groups of the species represented in `homerange`. Make a table the shows the counts for each of these `taxon`.**  
```{r}
table(homerange$taxon)
```

**6. The species in `homerange` are also classified into trophic guilds. How many species are represented in each trophic guild.**  
```{r}

trophoic_species<-select(homerange,trophic.guild,species)
trophoic_species
x<-unique(trophoic_species$species)
length(x)
```
The two trophic guilds are herbivore and carnivore, with 227 herbivore species and 342 carnivore species


# not due, but decent practice



**7. Make two new data frames, one which is restricted to carnivores and another that is restricted to herbivores.**  
```{r}
herbivore<-filter(homerange,trophic.guild!="carnivore")
herbivore

```


```{r}
carnivore<-filter(homerange,trophic.guild=="carnivore")
carnivore

```
**8. Do herbivores or carnivores have, on average, a larger `mean.hra.m2`? Remove any NAs from the data.**  
```{r}
meanc<-table(carnivore$mean.hra.m2)
mean(meanc, na.rm=TRUE)

```

```{r}
meanh<-table(herbivore$mean.hra.m2)
mean(meanh,na.rm=TRUE)
```
ON AVG carnivore has larger hra m2
**9. Make a new dataframe `owls` that is limited to the mean mass, log10 mass, family, genus, and species of owls in the database. Which is the smallest owl? What is its common name? Do a little bit of searching online to see what you can learn about this species and provide a link below** 
```{r}
owls<-select(homerange,mean.mass.g,log10.mass,family,genus,species)
owl<-filter(owls, family=="strigidae")

owl
```

```{r}
min(owl$mean.mass.g)
filter(owl,mean.mass.g==61.32)

```
the smallest owl is the Eurasian pygmy owl which is its common name. https://en.wikipedia.org/wiki/Eurasian_pygmy_owl. It is part of the strigidae family,glaucidium genus and passerinum species.
https://animaldiversity.org/accounts/Strigidae/ tells us that this includes every owl except one. They are known has typical owls, or true owls.
**10. As measured by the data, which bird species has the largest homerange? Show all of your work, please. Look this species up online and tell me about it!**.  
```{r}
birds<-filter(homerange, taxon=="birds")
birds
hmr<-max(birds$mean.hra.m2)
hmr
highest_hmr<-filter(birds,mean.hra.m2==241000000.00)
highest_hmr
```
The crested Caracara seems to have the largest homerange. Although listed as a carnivore, it is actually an omnivore found mostly in south America. it is a forager and is pretty common. Despite this, some do exist in mexico. It looks like a hawk and has sharp beaks and talons. It has long bright corlored legs with a black cap against its white neck . It actually walks a lot more than it flies!

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   