---
title: "Lab 3 Homework"
author: "Hugo Mahatdejkul"
date: "`r Sys.Date()`"
output:
  html_document: 
    theme: spacelab
    keep_md: yes
---

## Instructions
Answer the following questions and complete the exercises in RMarkdown. Please embed all of your code and push your final work to your repository. Your final lab report should be organized, clean, and run free from errors. Remember, you must remove the `#` for the included code chunks to run. Be sure to add your name to the author header above.  

Make sure to use the formatting conventions of RMarkdown to make your report neat and clean!  

## Load the tidyverse
```{r message=FALSE, warning=FALSE}
library(tidyverse)
```

### Mammals Sleep  
1. For this assignment, we are going to use built-in data on mammal sleep patterns. From which publication are these data taken from? Since the data are built-in you can use the help function in R. The name of the data is `msleep`.  
```{r}
msleep
```

The dataset of mammal sleep patterns is taken from V.M. Savage and G.B. West. It is from the national acadmy of sciences, 104 (3):1051-1056,2007

2. Store these data into a new data frame `sleep`.  
```{r}
sleep<-data.frame(msleep)
sleep
```

3. What are the dimensions of this data frame (variables and observations)? How do you know? Please show the *code* that you used to determine this below.  
```{r}
?dim
dim(sleep)
```

11 variables and 83 observations

4. Are there any NAs in the data? How did you determine this? Please show your code.  
```{r}
summary(sleep)
```


Yes there are NAs in the data. Using summary command tells me that there are NAs in some of the columns.

5. Show a list of the column names is this data frame.
```{r}
names(sleep)
?names()

```

6. How many herbivores are represented in the data?  
```{r}
f<-filter(sleep,vore=="herbi")
f #shows 32 rows.
table(sleep$vore)
```
there are 32 herbivores

7. We are interested in two groups; small and large mammals. Let's define small as less than or equal to 19kg body weight and large as greater than or equal to 200kg body weight. Make two new dataframes (large and small) based on these parameters.
```{r}
large<-filter(sleep,bodywt>=200)
small<-filter(sleep,bodywt<=19)
large
small
```

```{r}
newdataframelarge<-data.frame(large)
newdataframelarge
```

```{r}
newdataframesmall<-data.frame(small)
newdataframesmall
```

8. What is the mean weight for both the small and large mammals?
```{r}
smallmean<-c(small$bodywt)
mean(smallmean)


```
The mean weight for small is 1.797847kg
```{r}
largemean<-c(large$bodywt)
mean(largemean)
```
The mean weight for large is 1747.071kg



9. Using a similar approach as above, do large or small animals sleep longer on average?  

```{r}
LargeSleep<-c(large$sleep_total)
mean(LargeSleep)
```

```{r}
SmallSleep<-c(small$sleep_total)
mean(SmallSleep)
```

Smaller animals sleep much longer on avg.

10. Which animal is the sleepiest among the entire dataframe?

```{r}
summary(sleep)
max(sleep$sleep_total)
sleeptimedata<-filter(sleep,sleep_total==19.9)
sleeptimedata
```
From the summary we see that 19.9 hours is the maximum number of sleep hours, which corresponds to the little brown bat.
Assuming sleepiness corresponds to amount of sleep, little brown bat is sleepiest.

## Push your final code to GitHub!
Please be sure that you check the `keep md` file in the knit preferences.   