---
title: "warmup 6"
output: html_document
date: "2024-01-30"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Warm-up
1. Load the `bison.csv` data.
```{r}
library(tidyverse)
library(janitor)
bison<- read_csv ("data/bison.csv")
```
2. What are the dimesions and structure of the data?
```{r}
dim(bison)
glimpse(bison)
summary(bison)
```
3. We are only interested in code, sex, weight, year of birth. Restrict the data to these variables and store the dataframe as a new object.
```{r}
interest_bison<-
  bison %>%
  select(data_code,animal_sex,animal_weight,animal_yob)
interest_bison

```

4. Pull out the animals born between 1980-1990.
```{r}
year_birth<-
  interest_bison %>%
  filter(animal_yob>=1980,animal_yob<=1990 )
year_birth
```

5. How many male and female bison are represented between 1980-1990?
```{r}
female<-
  year_birth %>%
  filter(animal_sex=="F")
dim(female)

male<-
  year_birth %>%
  filter(animal_sex=="M")
dim(male)


```
21 male 414 female
6. Between 1980-1990, were males or females larger on average?
```{r}
mean(female$animal_weight)
mean(male$animal_weight)
```
male larger




